Network Configuration Synthesis with Abstract Topologies

### 2016-11-12 paper/abstraction/paper.tex
edits for 1-3

\textbf{Abstract---}
We develop techniques for automated reasoning about end-to-end properties of abstract network topologies. We develop a network synthesis tool, \sysname, to allow network operators to write a high-level routing policy over an abstraction of their network topology. From such a specification, \sysname synthesizes a collection of router configurations running the distributed Border Gateway Protocol (BGP) that are guaranteed to implement the centralized routing policy for any concrete instantiation of the abstract topology, and for any combination of network failures in the resulting concrete topology.
%
Additionally, properties proven by \sysname over the abstract topology, such as reachability between a source and destination under k-failures, are also guaranteed to hold under any concrete instantiation of the abstract topology.
%
Something something, expandable/incremental networks. Something something scales compilation to large networks. Something something, templates easier to understand.
%These strong guarantees allow operators to reason about the routing behavior of their network at a high-level abstraction, while also generating an efficient low-level implementation.
%
%In addition, operators are able to safely change or expand their network in any way, so long as the new resulting topology also respects the abstraction.
%
We test \sysname out on ... blah blah blah

@@

We present \sysname, a system to generate provably-correct role templates. It allows operators to input \emph{abstract network topologies} in terms of roles and their connectivity. For instance, they may specify roles for ``top-of-rack'' and ``aggregation" routers and specify that every top-of-rack router connects to at least two aggregation routers (to tolerate the failure of a link to an aggregation router). Other inputs to \sysname include the routing policy of the network and its fault-tolerance requirements (e.g., the number of link failures to tolerate). For routing policy, we borrow notation from Propane~\cite{propane},  which specifies traffic paths at a high-level of abstraction. But while Propane policies
refer concrete devices, \sysname policies refer to abstract roles.

Based on these inputs, the \sysname compiler generates one template per role. It guarantees that, when the templates are instantiated to configurations for {\em any} concrete topology that complies with the abstract topology, the result network will correctly implement the routing policy and meet the fault-tolerance requirements. We show how fault-tolerance analysis can be done over abstract topology itself using \todo{say something cool about about the inference rules and such}.

\sysname templates are also evolution-friendly. When the network migrates from one concrete topology to another, both of which comply with the abstract topology, it is guaranteed that only the configurations of devices that acquire or lose a neighboring device needs to change (e.g., to activate a port on which the new device is connected).  No other device in the network is impacted.

\todo{Talk about evaluation result here ... roll the point about faster compilation there.}
As an aside, a pleasant side effect of this new compiler
infrastructure, designed to make life easier on network operators, is
that for structured networks, it scales substantially better than its
competitor \propane does: as the number of individual devices in a
network grows, the number of distinct roles will often stay constant.
Consequently, the \sysname compiler may be two orders of magnitude
faster than \propane for large structured networks.

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, the contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation algorithms that support network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates
configuration templates and is up to two orders of magnitude faster
than its closest competitor on large networks.
\end{itemize}

@@

Route announcements contain additional attributes. A common one is {\em AS path}, which denotes the sequence of Autonomous Systems that will be traversed to reach the destination. When BGP is used inside a datacenter, each router tends to be its own AS. Optional attributes include one or more {\em communities}, which are arbitrary strings. Operators agree on what each community value means out of band.

A BGP router may receive one or more announcements from different neighbors for the same prefix. It selects the best one(s) based on its {\em import policy}. These policies are a function of the AS path, community values, or the neighbor that sent the announcement, and may even filter some announcements altogether (i.e., not use them even if no other announcements to the destination are available). Then, based on the {\em export policy}, the router announces the selected route to one or more neighbors, after adding itself to the AS path and, optionally, modifying the set of community values.

@@

We demonstrate the difficulty of configuring networks using the example data center network in Figure~\ref{fig:example}. The boxes denote routers. Using terminology for fat tree networks~\cite{fattree}, S[1--2] denote spine routers, A[1--8] denote aggregation routers, and T[1--8] denote top-of-rack (ToR) routers. The spine routers are connected to the
rest of the Internet through Peer[1--2]. The aggregation routers
bridge the spine to the ToR routers. The ToR routers attach to a set of servers (``a rack'') that host services with address prefixes P[1--8].

@@

The intended policy for this network is:
$p1)$ complete internal connectivity, \IE, all routers should be able to talk to each other;
$p2)$ services in Pods[1--2] should be globally accessible;
$p3)$ prefixes for global service should be aggregated into a covering prefix PG when announced outside;
$p4)$ services in Pods[3--4] should not be externally accessible;
$p5)$ prefer Peer 1 over Peer 2, \IE, when both peers announce a prefix, send traffic through peer 1;
$p6)$ datacenter routers should not transit traffic between peers 1 and 2; and
$p7)$ no loss in connectivity after any single-link failure.

@@

To simplify network configuration, the operations community is adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration file per router, operators author one configuration template per role.
%
A {\em role} refers to specific functionality and is played by one or more routers.\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor-specific configuration languages.}
%
The network in Figure~\ref{fig:example} has five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Templates have parameters for various aspects of the configuration (\EG, interface addresses, neighbor list) and are compiled to device configurations by instantiating the parameters using a database of network information.

@@

\para{Templates today} While templates help avoid a class of configuration errors (e.g., inconsistent routing adjacencies), they are difficult to author. Operators must manually decompose complex, network-wide routing policy into role templates such that the collective behavior of the configurations generated from them will correctly implement the policy, under failures and as the network evolves.

This task is error-prone, and no tools exist today for template verification; in fact, such tools would be difficult to build as templates have no semantics, of policies or topologies they were intended to work for, independent of the configurations they produce. Today, operators test (not verify) templates by simulating their generated configurations over the current or known future topologies.

@@

Such simulations provide some assurance that the templates work correctly for the simulated topologies but little assurance for other topologies that may arise after the network evolves. Templates that work for one topology may not work for seemingly-inconsequential variations.  Unfotunately, currently,
there is no way to predict which variations will behave correctly for a given set of templates.
Worse, when operators discover their old templates no longer work, a change to a template will
cause a change to all devices that use that template---an unacceptable disruption in many cases.  When that happens, network operators may abandon templates entirely and revert to hand-crafting configurations, using local patches to accommodate the topological change.
Such patches reintroduce the complexity and errors that templates were meant to prevent.

@@

Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}; it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both topologies, but that is not true. For instance, consider templates that disallow ``valley" paths, which go up, down, and then up again; it is a common to disallow valley paths in a data center because they are longer than shortest paths. These templates will work perfectly for Figure~\ref{fig:example} but silently violate the fault tolerance requirement ($p7$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-induced blackhole (\S\ref{sec:motivation}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces  the covering prefix PG (because it gets routes for P[3--4]). In contrast, in Figure~\ref{fig:example}, no such blackhole will occur after a single link failure because spine routers have two links to each pod.

>>>>

\textbf{Abstract---}
We develop techniques for automated reasoning about end-to-end properties of abstract network topologies. We develop a network synthesis tool, \sysname, to allow network operators to write a high-level routing policy over an abstraction of their network topology. From such a specification, \sysname synthesizes a collection of router configurations running the distributed Border Gateway Protocol (BGP) that are guaranteed to implement the centralized routing policy for any concrete instantiation of the abstract topology, and for any combination of network failures in the resulting concrete topology.
%
Additionally, properties proven by \sysname over the abstract topology, such as reachability between a source and destination under k-failures, are also guaranteed to hold under any concrete instantiation of the abstract topology.
%
%These strong guarantees allow operators to reason about the routing behavior of their network at a high-level abstraction, while also generating an efficient low-level implementation.
%
%In addition, operators are able to safely change or expand their network in any way, so long as the new resulting topology also respects the abstraction.

@@

We present \sysname, a system to generate provably-correct configuration templates. It allows operators to input \emph{abstract network topologies} in terms of roles and their connectivity. For instance, they may specify roles for ``top-of-rack'' and ``aggregation" routers and specify that every top-of-rack router connects to at least two aggregation routers (to tolerate the failure of a link to an aggregation router). Other inputs to \sysname include the routing policy of the network and its fault-tolerance requirements (e.g., the number of link failures to tolerate). For routing policy, we borrow notation from Propane~\cite{propane}, a language for specifying traffic paths at a high-level of abstraction. But while Propane policies
refer concrete devices, \sysname policies refer to abstract roles.

Based on these inputs, the \sysname compiler generates one template per role. It guarantees that, when the templates are instantiated to configurations for {\em any} concrete topology that respects the abstraction, the resulting network will correctly implement the routing policy and meet the fault-tolerance requirements. To check fault-tolerance properties, we develop a new analysis over the abstraction based on a collection of inference rules about the number of disjoint paths between devices in different roles.

\sysname templates are also evolution-friendly. When the network migrates from one concrete topology to another, both of which comply with the abstract topology, it is guaranteed that only the configurations of devices that acquire a new neighboring device need to change (e.g., to activate a port on which the new device is connected).  No other device in the network is impacted.

In addition to creating reusable templates, 
%synthesizing configurations by working directly with an abstract topology means that 
\sysname can scale substantially better than its nearest competitor \propane does: as the number of individual devices in a
network grows, the number of distinct roles will often stay constant.
Consequently, the \sysname compiler can be two orders of magnitude
faster than \propane, taking less than 10 seconds to synthesize templates for large structured networks.

%\todo{Talk about evaluation result here ... roll the point about faster compilation there.}

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, the contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation strategies that support network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates
configuration templates and is up to two orders of magnitude faster
than its closest competitor on large networks.
\end{itemize}

@@

Route announcements contain additional attributes. A common one is {\em AS path}, which denotes the sequence of Autonomous Systems that will be traversed to reach the destination. 
%When BGP is used inside a datacenter, each router tends to be its own AS. 
Optional attributes include one or more {\em communities}, which are arbitrary strings. Operators agree on what each community value means out of band.

A BGP router may receive one or more announcements from different neighbors for the same prefix. It selects the best one based on its local {\em import policy}. These policies are a function of the AS path, communities, or the neighbor that sent the announcement, and may even filter some announcements altogether (i.e., not use them even if no other announcements to the destination are available). Then, based on the local {\em export policy}, the router announces the selected route to one or more neighbors, after adding itself to the AS path and, optionally, modifying the set of communities.

@@

We demonstrate the difficulty of configuring networks using the example data center network in Figure~\ref{fig:example}. The boxes denote routers. Using terminology for fat tree networks~\cite{fattree}, S[1--2] denote spine routers, A[1--8] denote aggregation routers, and T[1--8] denote top-of-rack (ToR) routers. The spine routers are connected to the
rest of the Internet through neighbors N[1--2]. The aggregation routers
bridge the spine to the ToR routers. The ToR routers attach to a set of servers (``a rack'') that host services with address prefixes P[1--8].

@@

The intended policy for this network is:
$(1)$ complete internal connectivity, \IE, all routers should be able to talk to each other;
$(2)$ services in Pods[1--2] should be globally accessible;
$(3)$ prefixes for global service should be aggregated into a covering prefix PG when announced outside;
$(4)$ services in Pods[3--4] should not be externally accessible;
$(5)$ paths should be valley-free, \IE, never go down then back up.
$(6)$ prefer neighbor N1 over N2, \IE, when both neighbors announce a prefix, send traffic through N1;
$(7)$ datacenter routers should not transit traffic between N1 and N2; and
$(8)$ no loss in connectivity after any single-link failure.

@@

To simplify network configuration, the operations community is adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration file per router, operators author one configuration template per role.
%
A {\em role} refers to specific functionality and is played by one or more routers.\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor-specific configuration languages.}
%
The network in Figure~\ref{fig:example} might have five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Templates have parameters for various aspects of the configuration (\EG, interface addresses, neighbor list) and are compiled to device configurations by instantiating the parameters using a database of network information.

@@

While templates help avoid a class of configuration errors (e.g., inconsistent routing adjacencies), they are difficult to author. Operators must manually decompose complex, network-wide routing policy into role templates such that the collective behavior of the configurations generated from them will correctly implement the policy, under failures and as the network evolves.

%This task is error-prone, and no tools exist today for template verification; in fact, such tools would be difficult to build as templates have no semantics, of policies or topologies they were intended to work for, independent of the configurations they produce. Today, operators test (not verify) templates by simulating their generated configurations over the current or known future topologies.

@@

%Such simulations provide some assurance that the templates work correctly for the simulated topologies but little assurance for other topologies that may arise after the network evolves.

Templates that work for one topology may not work for seemingly-inconsequential variations.  
%Unfotunately, currently, there is no way to predict which variations will behave correctly for a given set of templates.
Worse, when operators discover their old templates no longer work, changing a template will
cause a change to all devices that use that template---an unacceptable disruption in many cases.  When that happens, operators may abandon templates entirely and revert to hand-crafting configurations, using local patches to accommodate the change.
Such patches reintroduce the complexity and errors that templates were meant to prevent.

@@

Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}; it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both topologies. However, because the templates are configured to disallow ``valley" paths, which go down, and then up, these templates will work perfectly for Figure~\ref{fig:example} but silently violate the fault tolerance requirement ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-induced blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). In contrast, in Figure~\ref{fig:example}, no such blackhole will occur after a single link failure because spine routers have two links to each pod.
